---
title: "How to use the tcG package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use the tcG package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction 

This package contains all the functions needed to fit the models presented in [Boutigny et al. (to be published)](link)^[Boutigny M, Ailliot P, Chaubet A, Naveau P, Saussol B (to be published). “Modelling rainfall from sub-hourly to daily scale with a heavy tailed meta-Gaussian model.” *Water Resources Research*.].

The main functions of the package are `tcG.fit`, `extGP.fit` that are used to fit the models and `res.plot` can be used to draw plots. You can check the documentation pages.

```{r setup}
library(tcG)
```

```{r,eval=F}
?tcG.fit
?extGP.fit
?res.plot

```


Two types of models can be fitted with this package, all aiming at modeling precipitation which we will note $Y$. 

### Meta-Gaussian models

A meta-Gaussian model can be written as

$$ Y = 0*\mathbb{I}_{X<0} + \psi(X)*\mathbb{I}_{X\geq0} \text{,       with } X\sim \mathcal{N}(\mu,1) $$
where $\mathbb{I}$ is the indicator function equals to 1 if the condition is true and 0 else. $\mu$ is the parameter that controls the probability of dry measurement.
The transformation $\psi$ is called the anamorphosis, and 4 options are available in the package:

- `gp` (for Generalized Pareto): $\psi(x) = y_m+\sigma x^{\frac{1}{\alpha}}\exp{\frac{\xi x^2}{2}} $
- `power`: $\psi(x) = y_m+\sigma x^{\frac{1}{\alpha}}$
- `quadratic-power`: $\psi(x) = y_m+\sigma_1 x^{\frac{1}{\alpha}}+\sigma_2 x^{\frac{2}{\alpha}}$ 
- `power-exp`: $\psi(x) = \sigma_2 (\exp(\sigma_1x^{1/\alpha})-1)$

$y_m$ is the minimal value that can be observed.


The choice of the anamorphosis will always be controlled by the argument `name`.


### Extended Generalized Pareto model

This model is the one of [Naveau et al. (2016)](https://agupubs.onlinelibrary.wiley.com/doi/pdfdirect/10.1002/2015WR018552)^[Naveau P, Huser R, Ribereau P, Hannart A (2016). “Modeling jointly low, moderate, and heavy rainfall intensities without a threshold selection.” *Water Resources Research*, **52**(4), 2753–2769.]. It  models only the positive part of the distribution and can be written as

$$ Y_+ =  y_m + \sigma H^{-1}_\xi(U^{1/\alpha})$$
where $U \sim Unif(0,1)$, $H_\xi$ is the cdf of a GPD and $y_m$ is the minimal value that can be observed.

## Data

Load the package and the provided data set. 

```{r}
#data(guip, package="tcG")

# using only October to avoid long calculations
guip=guip[lubridate::month(guip$date)==10,]

plot(guip[1:7000,], type="l")
```


## Fitting meta-Gaussian models

### First fit

As a first example the `gp` transformation will be fitted. The argument `name` is a vector containing the names of the transformations that I want to fit. `init` is a list, named according to those transformations, which contains initial values for the parameters:

```{r, eval=F}
init=list("power"=c(mu, sigma, alpha),
          "gp"=c(mu, sigma, alpha, xi),
          "power-exp"=c(mu, sigma1, sigma2, alpha),
          "quadratic-power"=c(mu, sigma1, sigma2, alpha))
```

$y_m$ is the minimal value that can be observed and $step$ is the precision of the data. Finally `R` and `bootstrap` control the bootstrap replicates.

```{r}
sort(unique(guip$R))

res=tcG.fit(guip$R, name="gp", init=list("gp"=c(-1,0.5,0.5,0.5)), ym=0.2, step=0.2, R=50)

# fitted parameters
res$par
```


The default results include a plot for low intensities with a barplot zoomed on the 20 first steps, and a quantile-quantile plot that gives a better view of the global fit. The light area on the qqplot give the 95\% interval computed with the bootstrap replicates.

### The impact of discretization

Taking into account the discretization of precipitation is especially important at fine time scales. In the package there are two options to fit the models: the continuous likelihood and the discrete one. Choosing one is made by inquiring the precision of the data. If `step==0` (default) the continuous likelihood is used, else it is the discrete one.
 
```{r}
sort(unique(guip$R))

res_cont=tcG.fit(guip$R, name="gp", init=list("gp"=c(-1,0.5,0.5,0.5)), ym=0, step=0, R=50)
res_dis=tcG.fit(guip$R, name="gp", init=list("gp"=c(-1,0.5,0.5,0.5)), ym=0.2, step=0.2, R=50)
```

### Comparing several meta-Gaussian models

```{r}
res=tcG.fit(guip$R, name=c("power", "gp"),
            init=list("power"=c(-1,1,2), "gp"=c(-1,.5,0.5,0.5)),
            ym=.2, step=.2, R=50)

# parameters
res$par

# AIC
which.min(res$AIC)
```


## Fitting the extended GP model

```{r}
# uniquement sur les mesures positives
res=extGP.fit(guip$R[guip$R>0], init=c(.5,.5,.5), ym=0.2, step=0.2, R=50)

res$par
```


## Comparing meta-Gaussian and extended GP

We're not going to use the default plots, but the `res.plot` function.

```{r}
res_tcG=tcG.fit(guip$R, name="gp", init=list("gp"=c(-1,.5,0.5,0.5)), ym=0.2, step=0.2, R=50, plots=FALSE)
res_eGP=extGP.fit(guip$R[guip$R>0], init=c(.5,0.5,0.5), ym=0.2, step=0.2, R=50, plots=FALSE)

# plot
res.plot(res.tcG=res_tcG, res.extGP=res_eGP, y=guip$R, ym=0.2, step=0.2, choice="qqplot")
```

